\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-07-29 paper.tex
Extended a lot the section on knowledge model
Completed the state of art regarding knowledge management

\begin{figure*}
        \centering
        \includegraphics[width=1.0\columnwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

Relying on RDF triples and Description Logics has advantages such as the
availability of numerous mature open-source libraries to manipulate the ontology,
interoperability with several major on-line knowledge bases (like {\sc
OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
guarantee of decidability (it is always possible to classify a Description
Logics ontology).

Tools to manipulate and reason over ontologies are readily available, mature,
and already well accepted in the robotic community~\cite{Tenorth2009a,
Lim2011}. While alternatives like \emph{Answer Set Programming} have also been
successfully investigated in robotics~\cite{Chen2010,Erdem2012}, in particular
to deal with non-monotonic reasoning, we did not actually hit any brick wall
while working with OWL ontologies. We may reconsider this choice at a later
stage, but until now it has proven an effective framework to quickly explore
implementations of new cognitive abilities (for instance, it has been
conceptually and technically easy to add support for independent knowledge
models, one per the agents the robot interacts with --see
section~\ref{sect|tom}).

@@

\paragraph{Knowledge model}

>>>>

\begin{figure*}
        \centering
        \includegraphics[width=0.8\textwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

At any time, the knowledge available to the robot may come from three sources:
Part of the knowledge is stored in an ontology (the {\sc OpenRobots Ontology}),
loaded when the robot initializes. This static source covers the
\emph{common-sense} knowledge of the robot, and optionally some
scenario-specific knowledge (for instance about objects that are to be
manipulated). The second part of the knowledge is acquired at runtime from
perception, interaction, decision making.  We go over the processes it involves
in the next section. Lastly, the third source of knowledge is the output of the
reasoner.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{top_tbox.pdf}

    \caption{The upper part of the ORO common-sense TBox. All these concepts
    belong to the {\sc OpenCyc} namespace.}

    \label{fig|upper_tbox}
\end{figure}

\paragraph{The OpenRobots Ontology} The ORO common-sense ontology has been designed from two requirements: covering
our experimental needs and conforming as much as possible to the {\sc OpenCyc}
upper ontology.

This lead to a bidirectional design process: from \emph{bottom-up} regarding
the choices of concepts to model, \emph{top-down} regarding the upper part of the
taxonomy. This upper part of the ontology is pictured on
figure~\ref{fig|upper_tbox}. All the classes visible on this figure belong to the
{\sc OpenCyc} namespace (the {\tt cyc:} prefix is omitted for clarity).

By \emph{aligning} the upper part of the ontology on {\sc OpenCyc} (as other
KRS, like {\sc KnowRob}~\cite{Tenorth2009a} or PEIS K\&R~\cite{Daoutis2009}, did)
has multiple advantages. First the design of this part of the ontology is
generally difficult: it pertains to abstract concepts whose mutual relations
comes to philosophical debates. The upper taxonomy of {\sc OpenCyc} represents
a relative consensus, at least within the semantic Web community. Then, because
it is a well established project with numerous links to other on-line databases
(like Wikipedia or WordNet), the reuse of important {\sc OpenCyc} concepts
ensures to a certain extend that the knowledge stored by the robot can be
shared or extended with well-defined semantics. A good example is the concept
of \emph{Object}: In everyday conversation, an object is a relatively small
physical thing, that can be typically manipulated. Normally, a human is not
considered as an object. In {\sc Cyc}, an object has a more precise definition:
it is something \emph{partially tangible}. That includes obviously the humans,
and actually many other entities that would not be commonly said to be objects
(the Earth for instance). Thus the importance of relying on well-defined and
standard semantics to exchange informations between artificial systems.

This figure~\ref{fig|upper_tbox} also illustrates the fundamental disjunction
in the ORO model between \emph{temporal} and \emph{spatial} entities (formally,
$(TemporalThing \sqcap SpatialThing)^{\mathcal{I}} = \emptyset$, with
$\mathcal{I}$ the \emph{interpretation} of our model).

The class \concept{purposeful action} is the superset of all the actions that
are voluntarily performed by the robot (or another agent). Subclasses (like
\concept{Give}, \concept{LookAt}, etc.) are not asserted in the common-sense
ontology, but are added by the execution controller (in link with the symbolic
task planner) and the natural language processor based on what is actually
performable and/or understandable by the robot at run-time.

The tree of figure~\ref{fig|upper_tbox} (this subset of the ontology is indeed
a tree: this has however not to be the case in general, and, as a matter of
fact, the TBox of the whole ORO common-sense ontology does not form a tree) is
not equally developed at lower levels. We have already briefly mentioned the
developments of the actions. The other important part is the descendants of the
\concept{partially tangible thing} (what is commonly called an \emph{object}).
Figure~\ref{fig|tangible_things_tbox} gives more details on this part of the
ontology.

\begin{figure}
    \centering
    \includegraphics[width=0.6\columnwidth]{tangible_things_tbox.pdf}
    \caption{TBox of the specialisations of \concept{PartiallyTangible}.}
    \label{fig|tangible_things_tbox}
\end{figure}

This excerpt from the ontology makes the bottom-up design process visible: only
few types of \emph{partially tangible} things appear, and only subclasses
relevant to the context of service robotics in an human-like environment are
present.

Lastly, the ORO common-sense ontology contains several rules and class
expressions that encode non-trivial inferences.

The definition of the \concept{Bottle} is a case in point:

\concept{Bottle} $\equiv$ \concept{Container} {\bf and} \concept{Tableware}
{\bf that} (\concept{hasShape} {\bf value} \concept{cylinderShape} {\bf and}
\concept{hasCharacteristicDimension} {\bf only} \concept{\em int[>= 0.1, <=
0.3]})

If a human informs the robot that a given object is indeed a bottle, the robot
can then infer much more on this object. And if the human affirms that a car is
a bottle, the robot may question this assertion because of the inconsistent
size.

We discuss in section~\ref{krs-discussion} the strengths and weaknesses of our
knowledge framework.

@@

\subsection{Knowledge framework}
\label{krs-discussion}

Knowledge flows in our architecture through a \emph{semantic blackboard}, the
{\sc Oro}-server. We rely on Description Logics (OWL) to represent and
manipulate facts.

Relying on RDF triples and Description Logics has decisive advantages such as
the availability of numerous mature open-source libraries to manipulate the
ontology, interoperability with several major on-line knowledge bases (besides
{\sc OpenCyc}, {\sc WordNet}, {\sc DBPedia} or {\sc
RoboEarth}~\cite{Waibel2011} are other example), open-world reasoning, and the
formal guarantee of decidability (it is always possible to classify a
Description Logics ontology).

Tools to manipulate and reason over ontologies are readily available, mature,
and already well accepted in the robotic community~\cite{Tenorth2009a,
Lim2011}. While alternatives like \emph{Answer Set Programming} have also been
successfully investigated in robotics~\cite{Chen2010,Erdem2012}, in particular
to deal with non-monotonic reasoning, we did not actually hit any brick wall
while working with OWL ontologies. We may reconsider this choice at a later
stage, but until now it has proven an effective framework to quickly explore
implementations of new cognitive abilities (for instance, it has been
conceptually and technically easy to add support for independent knowledge
models, one per agent the robot interacts with --see section~\ref{sect|tom}).
